Exploring probability distributions for bivariate temporal granularities

### 2019-12-08 main.Rmd
single to multiple all circular

# Linear Time granularities {#sec:lineartime}

@@

# Cyclic Time granularities {#sec:cyclic-gran-def}

@@

The cyclic calendar algebra consists of two kinds of operations: one that requires computing multiple order-up from single order-up cyclic granularities and the other that require the reverse. The hierarchical structure of time creates a natural nested ordering which can produce **single-order-up** or **multiple-order-up** granularities. We refer to granularities which are nested within multiple levels as **multiple-order-up** granularities and those concerning a single level as **single-order-up** granularities. We shall use the notion of a hierarchy table and order to define them.

**Notation:  Hierarchy table** Let $H_n: (G, C, P)$ be a hierarchy model containing n linear granularities. $G_{l}$ represents the linear granularity of order l and $P(l,m)$ represents the period length of $G_{l}$ with respect to $G_{m}$ and $C_{G(l),G(m)}$ represents the cyclic granularity that relates linear granularity of order l and m, $\forall l,m \in n, l<m$.  **Order** of a linear granularity can be comprehended as the level of graininess associated with a linear granularity. For example, if we consider two linear granularities $G$ and $H$, such that $G$ is finer than or groups into $H$, then $H$ is of higher order than $G$. In any hierarchy table, linear granularities are arranged from lowest to highest order of linear granularities. Let us look at some examples of single and multiple order-up granularities.

@@

<!-- some discussions on primary units of calendar -->
**Example:** So far we have used the Gregorian calendar as it is the most widely used calendar. But it is far from being the only one. All calendars fall under three types - solar, lunar or lunisolar/solilunar but the day is the basic unit of time underlying all calendars. Various calendars, however, use different conventions to structure days into larger units: weeks, months, years and cycle of years. The French revolutionary calendar divided each day into 10 "hours", each "hour" into 100 "minutes" and each "minute" into 100 "seconds". Nevertheless, for any calendar a hierarchy can be defined. For example, in Mayan calendar, one day was referred to as 1 kin and the calendar was structured as follows [@Reingold2001-kf]:

@@

## Multiple-order from single-order {#sec:multiplefromsingle}

\begin{equation} \label{eq7}
\begin{split}
C_{(G_l,G_m)}(z) & = C_{G_l,G_{l+1}}(z) + k(l, l+1)(C_{(G_(l+1),G_m)}(z)-1) \\  
  & =   C_{(G_l,G_{l+1}}(z) + k(l, l+1)[ C_{G_{l+1},G_{l+2}}(z) + k(l+1, l+2)( C_{G_{l+2}, G_m}(z) - 1) - 1] \\
  & =  C_{(G_l,G_{l+1}}(z) + k(l, l+1)(C_{G_{l+1},G_{l+2}}(z) - 1) + k(l, l+1)k(l+1, l+2)(C_{G_{l+2}, G_{m}} - 1)\\
  & =  C_{(G_l,G_{l+1}}(z) + k(l, l+1)(C_{G_{l+1},G_{l+2}}(z) - 1) + k(l, l+2)(C_{G_{l+2},G_{l+m}}(z) - 1)\\
  &\vdots\\
  & = \sum_{i=0}^{m - l - 1} k(l, l+i)(C_{G_{l+i},G_{l+i+1}}(z) - 1)\\
\end{split}
\end{equation}

Let us use the equation \ref{eq:eq3} to compute the multiple-order-up granularity uinal_katun for Mayan calendar, which is periodic and regular.

@@

## Single-order from multiple-order {#sec:singlefrommultiple}

>>>>

# Linear time granularities {#sec:lineartime}

@@

# Cyclic time granularities {#sec:cyclic-gran-def}

@@

The cyclic calendar algebra consists of two kinds of operations:  
(1) **single-to-multiple:** representation of multiple-order-up cyclic granularities from single order-up cyclic granularities  
(2) **multiple-to-single:** representation of single-order-up cyclic granularities from multiple order-up cyclic granularities

The hierarchical structure of time creates a natural nested ordering which can produce **single-order-up** or **multiple-order-up** granularities. We shall use the notion of a hierarchy table and order to define them.

**Hierarchy table** Let $H_n: (G, C, P)$ be a hierarchy model containing n linear granularities. $G_{l}$ represents the linear granularity of order l and $P(l,m)$ represents the period length of $G_{l}$ with respect to $G_{m}$ and $C_{G(l),G(m)}$ represents the cyclic granularity that relates linear granularity of order l and m, $\forall l,m \in n, l<m$.

**Order** of a linear granularity can be comprehended as the level of graininess associated with a linear granularity. For example, if we consider two linear granularities $G$ and $H$, such that $G$ is finer than or groups into $H$, then $H$ is of higher order than $G$. In any hierarchy table, linear granularities are arranged from lowest to highest order of linear granularities.

We refer to granularities which are nested within multiple levels of the hierarchy table as multiple-order-up granularities and those concerning a single level as single-order-up granularities.

@@

<!-- some discussions on primary units of calendar -->
**Example:** So far we have used example of cyclic granularities from Gregorian calendar as it is the most widely used calendar. But it is far from being the only one. All calendars fall under three types - solar, lunar or lunisolar/solilunar but the day is the basic unit of time underlying all calendars [@Reingold2001-kf]. Various calendars, however, use different conventions to structure days into larger units: weeks, months, years and cycle of years. The French revolutionary calendar divided each day into 10 "hours", each "hour" into 100 "minutes" and each "minute" into 100 "seconds". Nevertheless, for any calendar a hierarchy can be defined. For example, in Mayan calendar, one day was referred to as 1 kin and the calendar was structured as follows:

@@

## Single-to-multiple {#sec:single-to-multiple}

### All circular single order-up granularities

Circular single-order-up granularities can be used recursively to obtain multiple order up circular granularity. Since, the operation requires the use of modular arithmetic, it is important that the label mapping of each circular single order-up granularity is an identity function, that is, $L(x) = x\quad \forall x$. The label mapping of the resultant multiple-order-up granularity can however be chosen arbitrarily, depending on the context.

\begin{equation} \label{eq7}
\begin{split}
C_{(G_l,G_m)}(z) & = L(C_{G_l,G_{l+1}}(z) + k(l, l+1)(C_{(G_(l+1),G_m)}(z)-1)) \\  
  & =   L(C_{(G_l,G_{l+1}}(z) + k(l, l+1)[ C_{G_{l+1},G_{l+2}}(z) + k(l+1, l+2)( C_{G_{l+2}, G_m}(z) - 1) - 1]) \\
  & =  L(C_{(G_l,G_{l+1}}(z) + k(l, l+1)(C_{G_{l+1},G_{l+2}}(z) - 1) + k(l, l+1)k(l+1, l+2)(C_{G_{l+2}, G_{m}} - 1))\\
  & =  L(C_{(G_l,G_{l+1}}(z) + k(l, l+1)(C_{G_{l+1},G_{l+2}}(z) - 1) + k(l, l+2)(C_{G_{l+2},G_{l+m}}(z) - 1))\\
  &\vdots\\
  & = L(\sum_{i=0}^{m - l - 1} k(l, l+i)(C_{G_{l+i},G_{l+i+1}}(z) - 1))\\
\end{split}
\end{equation}

*Examplle*
Let us use the equation \ref{eq:eq3} to compute the multiple-order-up granularity uinal_katun for Mayan calendar, which is periodic and regular.

@@

### Circular or aperiodic single order-up granularities

## multiple-to-single {#sec:multiple-to-single}
